---
title: 'Future Hendrix: A Lyrical Analysis'
author: "Anthony Baker"
date: "5/28/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# For data manipulation
library(dplyr)
library(lubridate)
library(tidyr)
# For text analysis
library(tidytext)
library(stringr)
# For databse connections
library(DBI)
library(RMariaDB)
library(yaml)
# For API data formatting
library(jsonlite)
# For plotting
library(ggplot2)
```

```{r}
# Read from the DB
creds = read_yaml("db/genius_db.yml")
con <- dbConnect(RMariaDB::MariaDB(), 
                 username = creds$username,
                 password = creds$password,
                 host = creds$host,
                 dbname = creds$database)
# dbListTables(con)

res <- dbSendQuery(con, "SELECT * FROM songs_v3")

songs <- dbFetch(res) %>%
  as.tbl() %>%
  mutate(primary_artist_id = as.integer(primary_artist_id))
dbClearResult(res)
```

# Intro and Data Inspection
I've set up a database that contains lyrics and additional metadata from all Future songs on Rap Genius. Future is `primary_artist_id == 2197`.

```{r}
# How many songs?
songs %>% 
  nrow()
 # Any duplicates?
songs %>% 
  pull(song_id) %>%
  anyDuplicated()
```

```{r}
# How many languages?
songs %>% 
  count(lyrics_language, sort = T)
```

```{r}
# How many different primary artists?
songs %>% 
  count(primary_artist_id, sort = T)
```

```{r}
# How many songs are "published"?
songs %>% 
  count(published)
```

- There are `r songs %>% nrow()` songs in the dataset.
- There are no duplicated songs in the dataset.
- There are songs in 9 different languages, but `r round(843/866 * 100, 2)`% are in English.
- Future is the primary artist on 407 songs, or `r round(407/866 * 100, 2)`% of the dataset. He is featured in the remaining 457 songs.
- Only 57 of the songs are marked as officially "published" according to Genius.


# Who are Future's collaborators?
### Which artists' songs is Future most commonly featured on?
```{r}
# Which artists' songs is Future most commonly featured on?
# Add artist name via the Genius API:
add_artist_name <- function(artist_id){
  names <- rep(NA, length(artist_id))
  i <- 1
  for(id in artist_id){
    print(paste0("Fetching artist ", id))
    artist <- fromJSON(paste0("https://genius.com/api/artists/", id))
    names[i] <- artist$response$artist$name
    i <- i + 1
  }
  names
}

songs %>%
  filter(primary_artist_id != 2197) %>%
  count(primary_artist_id, sort = T) %>%
  head(10) %>%
  mutate(primary_artist_id = as.integer(primary_artist_id),
         name = add_artist_name(primary_artist_id))
```

### Which artists are most commonly featured on Future's songs?
```{r}
songs %>%
  filter(primary_artist_id == 2197) %>%
  select(featured_artist_ids) %>%
  mutate(featured_artist_ids = str_replace_all(featured_artist_ids, '\\[', ''), # Remove opening bracket in list
         featured_artist_ids = str_replace_all(featured_artist_ids, '\\]', ''), # Remove closing bracket in list
         featured_artist_ids = str_split(featured_artist_ids, ", ")) %>%
  unnest(featured_artist_ids) %>%
  filter(featured_artist_ids != "null") %>%
  count(featured_artist_ids, sort = T) %>%
  head(10) %>%
  mutate(name = add_artist_name(featured_artist_ids))
```

### Which artists are Future's most frequent collaborators, regardless of whose song it is?
```{r}
songs %>%
  select(song_id, primary_artist_id, featured_artist_ids) %>%
  mutate(featured_artist_ids = str_replace_all(featured_artist_ids, '\\[', ''), # Remove opening bracket in list
         featured_artist_ids = str_replace_all(featured_artist_ids, '\\]', ''), # Remove closing bracket in list
         featured_artist_ids = str_c(featured_artist_ids, ", ", primary_artist_id),
         featured_artist_ids = str_split(featured_artist_ids, ", ")) %>%
  unnest(featured_artist_ids) %>%
  filter(featured_artist_ids != "null",
         featured_artist_ids != 2197) %>%
  distinct(song_id, featured_artist_ids) %>%
  count(featured_artist_ids, sort = T) %>%
  head(10) %>%
  mutate(name = add_artist_name(featured_artist_ids))
```

### Which producers does Future work with most?
```{r}
songs %>% 
  filter(primary_artist_id == 2197) %>%
  # Some songs have multiple producers
  mutate(produced_by_temp = str_split(produced_by, "\\},")) %>%
  unnest(produced_by_temp) %>%
  mutate(producer_name = str_match(produced_by_temp, '"title": "(.+)",')[,2]) %>%
  select(song_id, produced_by, producer_name) %>%
  filter(!is.na(producer_name)) %>%
  distinct(song_id, producer_name) %>%
  count(producer_name, sort = T) %>%
  head(10)
```


***

# What are the most commonly used words?


```{r}
# Look at lyrics by line
future_lyrics <- songs %>%
  select(song_id, primary_artist_id, title, url, release_date, lyrics_language, lyrics) %>%
  mutate(release_date = as.Date(release_date),
         lyrics = str_replace(lyrics, '^\\["', ''), # Remove opening bracket in list
         lyrics = str_replace(lyrics, '\\"]$', ''), # Remove closing bracket in list
         lyrics = str_split(lyrics, '", "')) %>%
  unnest(lyrics) %>% 
  filter(lyrics != '') %>% # Don't include newlines
  # Create a verse artist column to distinguish who is rapping
  mutate(verse_artist = ifelse(str_detect(lyrics, '^\\[.*\\]$'), # Look for [Verse n: Artist]
                               lyrics,
                               NA),
         verse_artist = ifelse(str_detect(lyrics, ': '),
                               str_replace_all(verse_artist, '(^.*: )|(\\])', ''),
                               NA)) %>% # Isolate the artist name
  # Expand verse_artist for all relevant rows 
  fill(verse_artist) %>%
  # Exclude the verse signifiers
  filter(!str_detect(lyrics, '^\\[.*\\]$'))
```

```{r}
# Now that we have a `verse_artist`, let's look at who they are
future_lyrics %>% 
  distinct(song_id, verse_artist) %>%
  count(verse_artist, sort = T) %>%
  View()

# We can see that sometimes the verse artist is actually multiple people,
#   so when we filter by `verse_artist` we'll want to search for "future".
```

```{r}
# Now that we have lyrics by line and verse, let's break it down by word:
data("stop_words")

future_words <- future_lyrics %>%
  unnest_tokens(word, lyrics) %>%
  anti_join(stop_words, by = "word")
  

future_words %>%
  filter(str_detect(verse_artist, regex("future", ignore_case = T))) %>%
  count(word, sort = T) %>% 
  filter(n > 300) %>% 
  ggplot(aes(reorder(word, n), n)) +
  geom_col() +
  coord_flip() +
  labs(title = "Future Hendrix's most frequently used words",
       x = "word")
```

### How does Future's vocab selection compare to his collaborators?
```{r}
# TODO: Scrape all lyrics for Future's most frequent collaborators then do the comparison.
future_words %>%
  mutate(verse_artist_category = ifelse(
    str_detect(verse_artist, regex("future", ignore_case = T)),
    "Future",
    "Other"
  )) %>%
  filter(!is.na(verse_artist_category)) %>%
  count(verse_artist_category, word) %>%
  group_by(verse_artist_category) %>%
  mutate(proportion = n / sum(n)) %>%
  select(-n) %>%
  spread(verse_artist_category, proportion) %>%
  # gather(artist, proportion, Future, Other) %>%
  ggplot(aes(Other, Future, color = abs(Future - Other))) + 
    geom_abline(color = "gray40", lty = 2) +
    geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) + 
    scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75")
```

